### 2020-04-17 docs/analyses/analyses_additional.Rmd
- initalize GitHubPages

---
title: "How Do Like and Dislike Buttons Affect Communication?"
shorttitle : "Additional Analyses"
output: 
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    toc_depth: 2
---

```{r setup, include=F}
knitr::opts_chunk$set(
  cache=T, comment=NA, warning=F, echo=F, error=F, 
  message=F, prompt=F, tidy=T, fig.height=4
  )
options(digits = 3, width = 120)
```

```{r r-setup, include=F}
# define packages
packages <- c("cowplot", "devtools", "faoutlier", "GGally", "kableExtra",
              "knitr", "lavaan", "magrittr", "MVN", "psych", "pwr", "quanteda",
              "semTools", "tidyverse")

# load packages
lapply(packages, library, character.only = TRUE, quietly = TRUE)
```

```{r Data-Wrangling, include=F}
# load workspace
load("data/workspace.rdata")
```

# VIF

In what follows, you can find estimations of variance inflation factors to gauge multicollinearity. Generally, values above 5 or even 10 are considered problematic. However, these are only rules of thumb, and multcollinearity can occour with lower values. Indeed, although the values reported below are not above 5, they are increased, suggesting that multicollinearity might be at play here, which the regular analyses also confirm.

```{r vif, include=F}
# Self-Efficacy
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 
grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
SE01_01 ~~ a*SE01_02
SE01_03 ~~ a*SE01_04
trust_community =~ TR01_02 + TR01_03 + TR01_04
trust_provider =~ TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12
trust_spec =~ trust_community + trust_provider

self_eff ~ pri_con + grats_gen + pri_delib + trust_spec

# Covariates
GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + TR01_02 + TR01_03 + TR01_04 + TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu
"
fit <- sem(model, data = d, estimator = "MLR", missing = "ML")
r_self_eff <- inspect(fit, what = "rsquare")["self_eff"] # extract rsquare
vif_self_eff <- 1 / (1 - r_self_eff) # compute vif

# Privacy Deliberation
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 
grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
SE01_01 ~~ a*SE01_02
SE01_03 ~~ a*SE01_04
trust_community =~ TR01_02 + TR01_03 + TR01_04
trust_provider =~ TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12
trust_spec =~ trust_community + trust_provider

pri_delib ~ self_eff + pri_con + grats_gen + trust_spec

# Covariates
GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + TR01_02 + TR01_03 + TR01_04 + TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu
"
fit <- sem(model, data = d, estimator = "MLR", missing = "ML")
r_pri_delib <- inspect(fit, what = "rsquare")["pri_delib"]
vif_pri_delib <- 1 / (1 - r_pri_delib)

## Privacy Concerns
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 
grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
SE01_01 ~~ a*SE01_02
SE01_03 ~~ a*SE01_04
trust_community =~ TR01_02 + TR01_03 + TR01_04
trust_provider =~ TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12
trust_spec =~ trust_community + trust_provider

pri_con ~ self_eff + pri_delib + grats_gen + trust_spec

# Covariates
GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + TR01_02 + TR01_03 + TR01_04 + TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu
"
fit <- sem(model, data = d, estimator = "MLR", missing = "ML")
r_pri_con <- inspect(fit, what = "rsquare")["pri_con"]
vif_pri_con <- 1 / (1 - r_pri_con)

# Gratifications
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 
grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
SE01_01 ~~ a*SE01_02
SE01_03 ~~ a*SE01_04
trust_community =~ TR01_02 + TR01_03 + TR01_04
trust_provider =~ TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12
trust_spec =~ trust_community + trust_provider

grats_gen ~ self_eff + pri_con + pri_delib + trust_spec

# Covariates
GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + TR01_02 + TR01_03 + TR01_04 + TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu
"
fit <- sem(model, data = d, estimator = "MLR", missing = "ML")
r_grats_gen <- inspect(fit, what = "rsquare")["grats_gen"]
vif_grats_gen <- 1 / (1 - r_grats_gen)

# Trust
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 
grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
SE01_01 ~~ a*SE01_02
SE01_03 ~~ a*SE01_04
trust_community =~ TR01_02 + TR01_03 + TR01_04
trust_provider =~ TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12
trust_spec =~ trust_community + trust_provider

trust_spec ~ self_eff + pri_con + grats_gen + pri_delib

# Covariates
GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + TR01_02 + TR01_03 + TR01_04 + TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu
"
fit <- sem(model, data = d, estimator = "MLR", missing = "ML")
r_trust_spec <- inspect(fit, what = "rsquare")["trust_spec"]
vif_trust_spec <- 1 / (1 - r_trust_spec)
```

```{r}
# Table
tibble("Gratifications" = vif_grats_gen, "Trust Specific" = vif_trust_spec, "Privacy Concerns" = vif_pri_con, "Privacy Deliberation" = vif_pri_delib, "Self-Efficacy" = vif_self_eff) %>% 
  kable() %>% 
  kable_styling("striped")
```

# Additional analyses

In what follows, you can find slightly different models that were also run.

## Privacy calculus
### Model 1

Building on the Model "Adapted", we now remove the variable `Trust General`.

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
grats_inf =~ GR01_01 + GR01_02 + GR01_03 
grats_rel =~ GR01_04 + GR01_05 + GR01_06 
grats_par =~ GR01_07 + GR01_08 + GR01_09
grats_ide =~ GR01_10 + GR01_11 + GR01_12 
grats_ext =~ GR01_13 + GR01_14 + GR01_15
grats_spec =~ grats_inf + grats_rel + grats_par + grats_ide + grats_ext
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04

self_dis_log ~ a1*pri_con + b1*grats_spec + c1*pri_delib + d1*self_eff

# Covariates
self_dis_log + GR01_01 + GR01_02 + GR01_03 + GR01_04 + GR01_05 + GR01_06 + GR01_07 + GR01_08 + GR01_09 + GR01_10 + GR01_11 + GR01_12 + GR01_13 + GR01_14 + GR01_15 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu
"
fit <- sem(model, data = d, estimator = "MLR", missing = "ML")
summary(fit, fit = TRUE, std = TRUE)
```

Interestingly, now almost all effects disappear.

### Model 2

Here, we combine both privacy measures into a single one. Likewise, we combine both gratifications and trust.

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
pri_cau =~ pri_con + pri_delib
grats_inf =~ GR01_01 + GR01_02 + GR01_03 
grats_rel =~ GR01_04 + GR01_05 + GR01_06 
grats_par =~ GR01_07 + GR01_08 + GR01_09
grats_ide =~ GR01_10 + GR01_11 + GR01_12 
grats_ext =~ GR01_13 + GR01_14 + GR01_15
trust_community =~ TR01_02 + TR01_03 + TR01_04
trust_provider =~ TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12
grats_meta =~ grats_inf + grats_rel + grats_par + grats_ide + grats_ext + trust_community + trust_provider
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
self_dis_log ~ a1*pri_cau + b1*grats_meta + c1*self_eff

# Covariates
self_dis_log + GR01_01 + GR01_02 + GR01_03 + GR01_04 + GR01_05 + GR01_06 + GR01_07 + GR01_08 + GR01_09 + GR01_10 + GR01_11 + GR01_12 + GR01_13 + GR01_14 + GR01_15 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + TR01_02 + TR01_03 + TR01_04 + TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu
"
fit <- sem(model, data = d, estimator = "MLR", missing = "ML")
summary(fit, fit = TRUE, std = TRUE)
```

### Model 3

Here, we combine both privacy measures into a single one. Likewise, we combine both gratifications and trust.

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
pri_cau =~ pri_con + pri_delib
grats_inf =~ GR01_01 + GR01_02 + GR01_03 
grats_rel =~ GR01_04 + GR01_05 + GR01_06 
grats_par =~ GR01_07 + GR01_08 + GR01_09
grats_ide =~ GR01_10 + GR01_11 + GR01_12 
grats_ext =~ GR01_13 + GR01_14 + GR01_15
trust_community =~ TR01_02 + TR01_03 + TR01_04
trust_provider =~ TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12
grats_meta =~ grats_inf + grats_rel + grats_par + grats_ide + grats_ext + trust_community + trust_provider
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
self_dis_log ~ a1*pri_cau + b1*grats_meta + c1*self_eff

# Covariates
self_dis_log + GR01_01 + GR01_02 + GR01_03 + GR01_04 + GR01_05 + GR01_06 + GR01_07 + GR01_08 + GR01_09 + GR01_10 + GR01_11 + GR01_12 + GR01_13 + GR01_14 + GR01_15 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + TR01_02 + TR01_03 + TR01_04 + TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu
"
fit <- sem(model, data = d, estimator = "MLR", missing = "ML")
summary(fit, fit = TRUE, std = TRUE)
```

### Model 4

We now use also delete self-efficacy.

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
pri_cau =~ pri_con + pri_delib
grats_inf =~ GR01_01 + GR01_02 + GR01_03 
grats_rel =~ GR01_04 + GR01_05 + GR01_06 
grats_par =~ GR01_07 + GR01_08 + GR01_09
grats_ide =~ GR01_10 + GR01_11 + GR01_12 
grats_ext =~ GR01_13 + GR01_14 + GR01_15
trust_community =~ TR01_02 + TR01_03 + TR01_04
trust_provider =~ TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12
grats_meta =~ grats_inf + grats_rel + grats_par + grats_ide + grats_ext + trust_community + trust_provider
self_dis_log ~ a1*pri_cau + b1*grats_meta

# Covariates
self_dis_log + GR01_01 + GR01_02 + GR01_03 + GR01_04 + GR01_05 + GR01_06 + GR01_07 + GR01_08 + GR01_09 + GR01_10 + GR01_11 + GR01_12 + GR01_13 + GR01_14 + GR01_15 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + TR01_02 + TR01_03 + TR01_04 + TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 ~ male + age + edu
"
fit <- sem(model, data = d, estimator = "MLR", missing = "ML")
summary(fit, fit = TRUE, std = TRUE)
```

### Model 5

Here, we exchange specific gratitudes for general gratitudes.

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 
grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05

self_dis_log ~ a1*pri_con + b1*grats_gen

# Covariates
self_dis_log + GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 ~ male + age + edu
"
fit <- sem(model, data = d, estimator = "MLR", missing = "ML")
summary(fit, fit = TRUE, std = TRUE)
```

## Effect of popularity cues
### Like vs. control

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07  
grats_inf =~ GR01_01 + GR01_02 + GR01_03 
grats_rel =~ GR01_04 + GR01_05 + GR01_06 
grats_par =~ GR01_07 + GR01_08 + GR01_09
grats_ide =~ GR01_10 + GR01_11 + GR01_12 
grats_ext =~ GR01_13 + GR01_14 + GR01_15
grats_spec =~ grats_inf + grats_rel + grats_par + grats_ide + grats_ext

pri_con ~ a1*lik_ctrl
grats_spec ~ a2*lik_ctrl
self_dis_log ~ b1*pri_con + b2*grats_spec + c1*lik_ctrl

# Covariates
self_dis_log + GR01_01 + GR01_02 + GR01_03 + GR01_04 + GR01_05 + GR01_06 + GR01_07 + GR01_08 + GR01_09 + GR01_10 + GR01_11 + GR01_12 + GR01_13 + GR01_14 + GR01_15 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 ~ male + age + edu
"
fit_lik_ctrl_simple <- sem(model = model, data = dplyr::filter(d, !is.na(lik_ctrl)), estimator = "MLR", missing = "ML")
summary(fit_lik_ctrl_simple, fit = TRUE, std = TRUE)
```

### Like & dislike vs. control

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07  
grats_inf =~ GR01_01 + GR01_02 + GR01_03 
grats_rel =~ GR01_04 + GR01_05 + GR01_06 
grats_par =~ GR01_07 + GR01_08 + GR01_09
grats_ide =~ GR01_10 + GR01_11 + GR01_12 
grats_ext =~ GR01_13 + GR01_14 + GR01_15
grats_spec =~ grats_inf + grats_rel + grats_par + grats_ide + grats_ext

pri_con + grats_spec ~ likdis_ctrl
self_dis_log ~ a1*pri_con + b1*grats_spec + likdis_ctrl

# Covariates
self_dis_log + GR01_01 + GR01_02 + GR01_03 + GR01_04 + GR01_05 + GR01_06 + GR01_07 + GR01_08 + GR01_09 + GR01_10 + GR01_11 + GR01_12 + GR01_13 + GR01_14 + GR01_15 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 ~ male + age + edu
"
fit_likdis_ctrl_simple <- sem(model = model, data = dplyr::filter(d, !is.na(likdis_ctrl)), estimator = "MLR", missing = "ML")
summary(fit_likdis_ctrl_simple, std = TRUE)
```

### Like & dislike vs. like

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07  
grats_inf =~ GR01_01 + GR01_02 + GR01_03 
grats_rel =~ GR01_04 + GR01_05 + GR01_06 
grats_par =~ GR01_07 + GR01_08 + GR01_09
grats_ide =~ GR01_10 + GR01_11 + GR01_12 
grats_ext =~ GR01_13 + GR01_14 + GR01_15
grats_spec =~ grats_inf + grats_rel + grats_par + grats_ide + grats_ext

pri_con + grats_spec ~ likdis_lik
self_dis_log ~ a1*pri_con + b1*grats_spec + likdis_lik

# Covariates
self_dis_log + GR01_01 + GR01_02 + GR01_03 + GR01_04 + GR01_05 + GR01_06 + GR01_07 + GR01_08 + GR01_09 + GR01_10 + GR01_11 + GR01_12 + GR01_13 + GR01_14 + GR01_15 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 ~ male + age + edu
"
fit_likdis_lik <- sem(model = model, data = dplyr::filter(d, !is.na(likdis_lik)), estimator = "MLR", missing = "ML")
summary(fit_likdis_lik, std = TRUE, fit = TRUE)
```

# Results without covariates

As stated in our preregistration, we also provide the results of all analyses without controlling for covariates. The results remain the same.

## Privacy calculus
### Model "Peregistered"

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 
grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
trust_community =~ TR01_02 + TR01_03 + TR01_04
trust_provider =~ TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12
trust_spec =~ trust_community + trust_provider

self_dis_log ~ a1*pri_con + b1*grats_gen + c1*pri_delib + d1*self_eff + e1*trust_spec
"
fit_prereg <- sem(model, data = d, estimator = "MLR", missing = "ML")
summary(fit_prereg, fit = TRUE, std = TRUE)
```

### Model "Adapted"

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
grats_inf =~ GR01_01 + GR01_02 + GR01_03 
grats_rel =~ GR01_04 + GR01_05 + GR01_06 
grats_par =~ GR01_07 + GR01_08 + GR01_09
grats_ide =~ GR01_10 + GR01_11 + GR01_12 
grats_ext =~ GR01_13 + GR01_14 + GR01_15
grats_spec =~ grats_inf + grats_rel + grats_par + grats_ide + grats_ext
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
trust_gen =~ TR01_01 + TR01_05 + TR01_09

self_dis_log ~ a1*pri_con + b1*grats_spec + c1*pri_delib + d1*self_eff + e1*trust_gen
"
fit_adapted <- sem(model, data = d, estimator = "MLR", missing = "ML", missing = "ML")
summary(fit_adapted, fit = TRUE, std = TRUE)
```

### Model "Simple"

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07  
grats_inf =~ GR01_01 + GR01_02 + GR01_03 
grats_rel =~ GR01_04 + GR01_05 + GR01_06 
grats_par =~ GR01_07 + GR01_08 + GR01_09
grats_ide =~ GR01_10 + GR01_11 + GR01_12 
grats_ext =~ GR01_13 + GR01_14 + GR01_15
grats_spec =~ grats_inf + grats_rel + grats_par + grats_ide + grats_ext

self_dis_log ~ a1*pri_con + b1*grats_spec
"
fit_simple <- sem(model, data = d, estimator = "MLR", missing = "ML")
summary(fit_simple, fit = TRUE, std = TRUE)
```

## Effects of popularity cues
### Preregistered
#### Like vs. Control

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
trust_spec =~ trust_spec_fs

pri_con + grats_gen + pri_delib + self_eff + trust_spec ~ lik_ctrl
self_dis_log ~ a*pri_con + b*grats_gen + c*pri_delib + d*self_eff + e*trust_spec + f*lik_ctrl
"
fit_lik_ctrl <- sem(model = model, data = dplyr::filter(d, !is.na(lik_ctrl)), estimator = "MLR", missing = "ML")
summary(fit_lik_ctrl, fit = TRUE, std = TRUE)
```

#### Like & Dislike vs. Control

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
trust_spec =~ trust_spec_fs

pri_con + grats_gen + pri_delib + self_eff + trust_spec ~ likdis_ctrl
self_dis_log ~ a*pri_con + b*grats_gen + c*pri_delib + d*self_eff + e*trust_spec + f*likdis_ctrl
"
fit_likdis_ctrl <- sem(model = model, data = dplyr::filter(d, !is.na(likdis_ctrl)), estimator = "MLR", missing = "ML")
summary(fit_likdis_ctrl, fit = TRUE, std = TRUE)
```

#### Like & Dislike vs. Like

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
trust_community =~ TR01_02 + TR01_03 + TR01_04
trust_provider =~ TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12
trust_spec =~ trust_community + trust_provider

pri_con + grats_gen + pri_delib + self_eff + trust_spec ~ likdis_lik
self_dis_log ~ a*pri_con + b*grats_gen + c*pri_delib + d*self_eff + e*trust_spec + f*likdis_lik
"

fit_likdis_lik <- sem(model = model, data = dplyr::filter(d, !is.na(likdis_lik)), estimator = "MLR", missing = "ML")
summary(fit_likdis_lik, fit = TRUE, std = TRUE)
```

# Results including removed participants

As stated in the preregistration, we also report the analyses including the deleted participants. Results don't change meaningfully.

## Privacy calculus
### Model "Peregistered"

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 
grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
trust_community =~ TR01_02 + TR01_03 + TR01_04
trust_provider =~ TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12
trust_spec =~ trust_community + trust_provider

self_dis_log ~ a1*pri_con + b1*grats_gen + c1*pri_delib + d1*self_eff + e1*trust_spec

# Covariates
self_dis_log + GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + TR01_02 + TR01_03 + TR01_04 + TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu
"
fit_prereg <- sem(model, data = d_all, estimator = "MLR", missing = "ML")
summary(fit_prereg, fit = TRUE, std = TRUE)
```

### Model "Adapted"

Building on the preregistered model, instead of general gratifications and specific trust, we now use specific gratifications and general trust.

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
grats_inf =~ GR01_01 + GR01_02 + GR01_03 
grats_rel =~ GR01_04 + GR01_05 + GR01_06 
grats_par =~ GR01_07 + GR01_08 + GR01_09
grats_ide =~ GR01_10 + GR01_11 + GR01_12 
grats_ext =~ GR01_13 + GR01_14 + GR01_15
grats_spec =~ grats_inf + grats_rel + grats_par + grats_ide + grats_ext
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
trust_gen =~ TR01_01 + TR01_05 + TR01_09

self_dis_log ~ a1*pri_con + b1*grats_spec + c1*pri_delib + d1*self_eff + e1*trust_gen

# Covariates
self_dis_log + GR01_01 + GR01_02 + GR01_03 + GR01_04 + GR01_05 + GR01_06 + GR01_07 + GR01_08 + GR01_09 + GR01_10 + GR01_11 + GR01_12 + GR01_13 + GR01_14 + GR01_15 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + TR01_01 + TR01_05 + TR01_09 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu
"
fit_adapted <- sem(model, data = d_all, estimator = "MLR", missing = "ML", missing = "ML")
summary(fit_adapted, fit = TRUE, std = TRUE)
```

### Model "Simple"

We now use only variables, that is specific gratifications and privacy concerns.

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07  
grats_inf =~ GR01_01 + GR01_02 + GR01_03 
grats_rel =~ GR01_04 + GR01_05 + GR01_06 
grats_par =~ GR01_07 + GR01_08 + GR01_09
grats_ide =~ GR01_10 + GR01_11 + GR01_12 
grats_ext =~ GR01_13 + GR01_14 + GR01_15
grats_spec =~ grats_inf + grats_rel + grats_par + grats_ide + grats_ext

self_dis_log ~ a1*pri_con + b1*grats_spec

# Covariates
self_dis_log + GR01_01 + GR01_02 + GR01_03 + GR01_04 + GR01_05 + GR01_06 + GR01_07 + GR01_08 + GR01_09 + GR01_10 + GR01_11 + GR01_12 + GR01_13 + GR01_14 + GR01_15 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 ~ male + age + edu
"
fit_simple <- sem(model, data = d_all, estimator = "MLR", missing = "ML")
summary(fit_simple, fit = TRUE, std = TRUE)
```

## Effects of popularity cues
### Preregistered

#### Like vs. Control

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
trust_spec =~ trust_spec_fs

pri_con + grats_gen + pri_delib + self_eff + trust_spec ~ lik_ctrl
self_dis_log ~ a*pri_con + b*grats_gen + c*pri_delib + d_all*self_eff + e*trust_spec + f*lik_ctrl

# Covariates
self_dis_log + GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu
"
fit_lik_ctrl <- sem(model = model, data = dplyr::filter(d_all, !is.na(lik_ctrl)), estimator = "MLR", missing = "ML")
summary(fit_lik_ctrl, fit = TRUE, std = TRUE)
```

#### Like & Dislike vs. Control

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
trust_spec =~ trust_spec_fs

pri_con + grats_gen + pri_delib + self_eff + trust_spec ~ likdis_ctrl
self_dis_log ~ a*pri_con + b*grats_gen + c*pri_delib + d_all*self_eff + e*trust_spec + f*likdis_ctrl

# Covariates
self_dis_log + GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu
"
fit_likdis_ctrl <- sem(model = model, data = dplyr::filter(d_all, !is.na(likdis_ctrl)), estimator = "MLR", missing = "ML")
summary(fit_likdis_ctrl, fit = TRUE, std = TRUE)
```

#### Like & Dislike vs. Like

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
trust_community =~ TR01_02 + TR01_03 + TR01_04
trust_provider =~ TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12
trust_spec =~ trust_community + trust_provider

pri_con + grats_gen + pri_delib + self_eff + trust_spec ~ likdis_lik
self_dis_log ~ a*pri_con + b*grats_gen + c*pri_delib + d_all*self_eff + e*trust_spec + f*likdis_lik

# Covariates
self_dis_log + GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + TR01_02 + TR01_03 + TR01_04 + TR01_06 + TR01_07 + TR01_08 + TR01_10 + TR01_11 + TR01_12 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu
"

fit_likdis_lik <- sem(model = model, data = dplyr::filter(d_all, !is.na(likdis_lik)), estimator = "MLR", missing = "ML")
summary(fit_likdis_lik, fit = TRUE, std = TRUE)
```