### 2020-04-17 docs/data/items.rmd

---
title: "How Do Like and Dislike Buttons Affect Communication?"
subtitle: "List of Items"
output: 
  pdf_document:
    toc: TRUE
    toc_depth: 3
---

\newpage

# Response format

All items were measured on a 7-point scale with Likert-response format. Answers ranged between:

 - &nbsp;-3   completely disagree
 - &nbsp;-2   disagree
 - &nbsp;-1   slightly disagree
 - &nbsp;&nbsp;0   neutral
 - +1   slightly agree
 - +2   agree
 - +3   strongly agree

All items were translated from German.

# Privacy concerns

Using the participation platform I had ...

 1. ... concerns about what happens to my data.
 2. ... concerns about disclosing information about myself.
 3. ... no concerns.
 4. ... concerns that others could discover my real identity (i.e. my last and first name).
 5. ... concerns that information about myself could fall into wrong hands.
 6. ... concerns that others could discover what my political views are.
 7. ... concerns about my privacy.

# Gratifications general

Using the participation platform ...

 1.	... had many benefits for me.
 2.	... has paid off for me.
 3.	... was worthwhile.
 4.	... was fun.
 5.	... has brought me further regarding content.

# Gratifications specific

Using the participation platform it has been possible for me ...

_Information_

 1.	... to learn things I would not otherwise have noticed.
 2.	... to hear the opinion of others.
 3.	... to learn how other people tick.

_Relevance_

 4.	... to react to a subject that is very dear to me.
 5.	... to react to a subject that is important to me.
 6.	... to react to a subject that I am affected by.

_Political participation_

 7.	... to engage politically.
 8.	... to discuss political issues.
 9.	... to pursue my political interest.

_Idealism_

 10. ... to try to improve society.
 11. ... to advocate to something meaningful.
 12. ... to serve a good purpose.

_Extrinsic benefits_

 13. ... to do the responsible persons a favor.
 14. ... to soothe my guilty consciences.
 15. ... to fulfil my civic duty.

# Privacy deliberation

Using the participation platform ...

 1. ... I have considered whether I could be disadvantaged by writing a comment.
 2. ... I have considered whether I could be advantaged by writing a comment.
 3. ... I have weighed up the advantages and disadvantages of writing a comment.
 4. ... I have thought about consequences of a possible comment.
 5. ... I have considered whether I should write a comment or not.

# Trust general

&nbsp;&nbsp;&nbsp;1. The other users seemed trustworthy.  
&nbsp;&nbsp;&nbsp;5. The operators of the participation platform seemed trustworthy.  
&nbsp;&nbsp;&nbsp;9. The website seemed trustworthy.

# Trust specific

_Community_

 2.	The comments of other users were useful.
 3.	The other users had good intentions.
 4.	I could rely on the statements of other users.

_Provider_

 6.	The operators of the participation platform have done a good job.
 7.	It was important to the operators that the users are satisfied with the participation platform.
 8.	I could rely on the statements of the operators of the participation platform.

_Information System_

 10.	The website worked well.
 11.	I had the impression that my data was necessary for the use of the website.
 12.	I found the website useful.

# Self-efficacy

 1. In principle, I felt able to write a comment.
 2. I felt technically competent enough to write a comment.
 3. In terms of the topic, I felt competent enough to express my opinion.
 4. I found the topic too difficult to express my opinion.
 5. I found it complicated to write a comment.
 6. I was overburdened to write a comment.

# Self-disclosure

Let us assume that the participation platform would continue to exist in the future. Please indicate to what extent you agree with the following statements:

In the future, I am willing to ...

 1. ... discuss political issues on the participation platform.
 2. ... express my personal opinion on the participation platform.
 3. ... write comments on the participation platform.
 4. ... disclose concerns and fears on the participation platform.
 5. ... share personal experience on the participation platform.