### 2020-04-17 index.Rmd
- updated header

---
title: "Privacy Calculus Experiment"
---

On this repository, you can find the data, analysis script, and a reproducible manuscript of the paper "How Do Like and Dislike Buttons Affect Communication? A Privacy Calculus Approach to Understanding Self-Disclosure Online in a One-Week Field Experiment".

To view the results / additional analyses, you can either
 - download all materials and open the file "analyses/analyses.html" in your browse
 - or simply click on <a href="https://htmlpreview.github.io/?https://github.com/tdienlin/privacy_calc_exp/blob/master/analyses/analyses.html" target="_blank">this link</a>, and the results will open up directly in your browser.

To rerun the analyses, please do the following steps:

 - download all materials
 - open the RStudio project (file "privacy_calc_exp.Rproj")
 - run the file "analyses.rmd"
 - please make sure that all packages are installed using their current version

To reproduce the paper, please do the following steps **in addition**

 - knit file "manuscript/manuscript.pdf"
 - make sure that a current version of latex is installed