### 2020-04-17 index.Rmd
- initialized webpage

---
title: "Privacy Calculus Experiment"
---

On this website, you can find the companion material for the paper "How Do Like and Dislike Buttons Affect Communication? A Privacy Calculus Approach to Understanding Self-Disclosure Online in a One-Week Field Experiment".

The material for example includes the [data](data/data.csv), [analysis script](analyses.html), and a reproducible [manuscript](manuscript.html) of the paper.

This website is produced directly from the project's [github repository](https://github.com/tdienlin/privacy_calc_exp). You can download the material or file an issue.